% interaction/security.tex — completeness, soundness, knowledge soundness, claim trees

\section{Security Definitions}\label{sec:interaction-security}

Security definitions are conceptually the same as in the flat-indexed model but
now sit on the $\mathsf{Spec} + \mathsf{RoleDecoration}$ foundation.  At the
oracle level, the key difference is that verifier-side acceptance is phrased in
terms of the \emph{existence} of concrete oracle data compatible with the
verifier's query-level simulation, rather than by assuming a built-in
reification function.

Thus the security layer changes much less than the interaction layer itself:
the semantics are richer, but the core completeness/soundness/knowledge
soundness pattern remains the same.

\subsection{Protocol participants}

\begin{definition}[Prover, Verifier, Reduction]
  \label{int:reduction}
  A \emph{prover} takes $(\mathit{stmt}, \mathit{wit})$ and produces, via
  monadic setup, a $\Strategy.\mathsf{withRoles}$ whose output is
  $\mathsf{HonestProverOutput}\;(\mathit{StmtOut}\;s\;\mathit{tr})\;
  (\mathit{WitOut}\;s\;\mathit{tr})$.
  A \emph{verifier} is a statement-indexed $\mathsf{Counterpart}$ with
  $\mathit{StmtOut}\;s\;\mathit{tr}$ at $\mathsf{done}$.
  A \emph{reduction} pairs a prover with a verifier.
  $\mathit{WitnessIn}$ is intentionally not statement-dependent; compatibility
  is expressed in the security relations.
  \lean{Interaction.Reduction}
  \uses{int:strategy-with-roles, int:counterpart}
\end{definition}

$\Reduction.\mathsf{execute}$ runs the prover's strategy against the verifier
via $\mathsf{runWithRoles}$, returning the transcript and both outputs.
$\Reduction.\mathsf{Continuation}$ supports transcript-indexed second-stage
composition.

\subsection{Completeness, soundness, knowledge soundness}

All definitions use a generic monad~$m$ with $[\mathsf{HasEvalSPMF}\;m]$ for
probability semantics.

\begin{definition}[Completeness]
  \label{int:completeness}
  A reduction satisfies \emph{completeness with error $\varepsilon$} if for all
  valid inputs, honest execution produces valid output with probability at
  least $1 - \varepsilon$.
  \lean{Interaction.Reduction.completeness}
  \uses{int:reduction}
\end{definition}

\begin{definition}[Soundness]
  \label{int:soundness}
  A verifier satisfies \emph{soundness with error $\varepsilon$} if for all
  inputs outside the input language and all (possibly malicious) provers, the
  probability that the verifier's output falls in the output language is at
  most~$\varepsilon$.  Thus soundness is verifier-side: the honest prover does
  not appear in the definition.
  \lean{Interaction.soundness}
  \uses{int:reduction}
\end{definition}

\begin{definition}[Knowledge Soundness]
  \label{int:knowledge-soundness}
  A verifier satisfies \emph{knowledge soundness with error $\varepsilon$} if
  there exists a straightline extractor such that for all provers, the
  probability that the verifier accepts but the extractor fails to produce a
  valid witness is at most~$\varepsilon$.
  \lean{Interaction.knowledgeSoundness}
  \uses{int:reduction}
\end{definition}

\subsection{Composition theorems}

\begin{theorem}[Completeness composes]
  \label{thm:completeness-comp}
  If reduction~1 has completeness error~$\varepsilon_1$ and reduction~2
  (a continuation) has completeness error~$\varepsilon_2$ at every first-phase
  transcript, then the composed reduction has completeness error at most
  $\varepsilon_1 + \varepsilon_2$.
  \lean{Interaction.Reduction.completeness_comp}
  \uses{int:completeness}
\end{theorem}

Analogous theorems hold for perfect completeness and soundness.

\subsection{Round-by-round analysis}

\begin{definition}[Claim Tree]
  \label{int:claim-tree}
  A \emph{claim tree} is a recursive soundness witness on $\mathsf{Spec} +
  \mathsf{RoleDecoration}$.  At each sender (prover-message) node, bad claims
  must stay bad.  At each receiver (verifier-challenge) node, a bad claim may
  flip to good with probability at most $\mathit{error}$.
  \lean{Interaction.ClaimTree}
  \uses{int:role-decoration}
\end{definition}

\begin{theorem}[Terminal probability bound]
  \label{thm:bound-terminal-prob}
  If a $\mathsf{ClaimTree}$ is sound with per-round error bounds, then the
  probability of reaching a good terminal claim from a bad root is bounded by
  the sum of the per-round errors along any path.
  \lean{Interaction.ClaimTree.IsSound.bound_terminalProb}
  \uses{int:claim-tree}
\end{theorem}

\subsection{Oracle security}\label{subsec:oracle-security}

Security definitions for oracle reductions involve three layers.
An \emph{input implementation} is a deterministic query implementation
$\mathit{InputImpl} := \mathsf{QueryImpl}\;[\mathit{OStmtIn}]_o\;\mathsf{Id}$,
concretizing the input oracles.
An \emph{output implementation} (or \emph{simulate}) is a monadic query
implementation $\mathsf{QueryImpl}\;[\mathit{OStmtOut}]_o\;(\mathsf{OracleComp}\;\ldots)$
that the verifier exposes for its output oracles, without committing to
concrete underlying data.
$\mathsf{OutputRealizes}$ connects the two: it asserts query-level agreement
between a simulate function and a concrete output oracle family
$\mathit{oStmtOut}$, relative to a deterministic input implementation
and the transcript.

\begin{definition}[Oracle input/output relation]
  \label{int:oracle-relation}
  An \emph{input relation} takes $(\mathit{shared}, \mathit{stmt},
  \mathit{inputImpl}, \mathit{witIn})$ and returns $\mathsf{Prop}$.
  An \emph{output relation} takes $(\mathit{shared}, \mathit{inputImpl},
  \mathit{pt}, \mathit{stmtOut}, \mathit{outputImpl}, \mathit{witOut})$
  and returns $\mathsf{Prop}$.
  An \emph{input/output language} is the witness-free variant.
  \lean{Interaction.Oracle.Reduction.InputRelation,
        Interaction.Oracle.Reduction.OutputRelation}
\end{definition}

\begin{definition}[Oracle completeness]
  \label{int:oracle-completeness}
  An oracle reduction satisfies \emph{completeness with error $\varepsilon$}
  if, for every valid input $(\mathit{shared}, s, w)$ with $\mathit{relIn}$
  satisfied, honest execution produces output satisfying three conditions
  with probability at least $1 - \varepsilon$:
  \begin{enumerate}
    \item The prover's output statement agrees with the verifier's.
    \item $\mathsf{OutputRealizes}$ holds: the verifier's simulate function
      agrees with the prover's concrete output oracle data
      $\mathit{oStmtOut}$.
    \item $\mathit{relOut}$ holds for the output.
  \end{enumerate}
  \lean{Interaction.Oracle.Reduction.completeness}
  \uses{int:oracle-relation}
\end{definition}

\begin{definition}[Oracle knowledge soundness]
  \label{int:oracle-knowledge-soundness}
  A verifier satisfies \emph{knowledge soundness with error $\varepsilon$}
  if there exists a straightline extractor~$E$ such that for every malicious
  prover, the following holds.
  The adversarial prover is required to output both concrete output oracle
  data $\mathit{oStmtOut}$ and an output witness $\mathit{witOut}$.
  The event of interest is:
  $\mathsf{OutputRealizes}(\mathit{simulate}, \mathit{oStmtOut})$
  $\wedge$ $\mathit{relOut}(\ldots, \mathit{simulate}, \mathit{witOut})$
  $\wedge$ $\neg\,\mathit{relIn}(\ldots,
    E(\ldots, \mathit{oStmtOut}, \mathit{simulate}, \mathit{witOut}))$.
  The probability of this event is at most~$\varepsilon$.

  The extractor sees the transcript, input statement, input implementation,
  output statement, concrete $\mathit{oStmtOut}$, simulate, and
  $\mathit{witOut}$, and produces a candidate $\mathit{witIn}$.
  \lean{Interaction.Oracle.Verifier.knowledgeSoundness}
  \uses{int:oracle-relation, int:oracle-completeness}
\end{definition}

\begin{definition}[Oracle soundness]
  \label{int:oracle-soundness}
  A verifier satisfies \emph{soundness with error $\varepsilon$} if for
  every input outside the input language and every malicious prover, the
  probability that the verifier's output falls in the output language is at
  most~$\varepsilon$.
  Soundness follows from knowledge soundness under standard hypotheses (the
  input language is the witness-free projection of the input relation, and
  the output language is implied by the output relation for any witness).
  \lean{Interaction.Oracle.Verifier.soundness}
  \uses{int:oracle-knowledge-soundness}
\end{definition}
