% Copyright (c) 2025 ZKLib Contributors. All rights reserved.
% Released under Apache 2.0 license as described in the file LICENSE.

\section{FRI}\label{sec:fri}

FRI (Fast Reed--Solomon Interactive Oracle Proof of Proximity) tests whether an oracle
$f : \evaldomain \to \field$ is close to a Reed--Solomon code.  The ArkLib formalization
follows the generalized FRI low-degree-test presentation in \cite{FRI1216} and the
proximity-gap and Batched-FRI analysis of \cite{BCIKS20}.  It uses the usual two-phase structure:
a folding phase, implemented as a sequential composition of single-round oracle reductions, and a
query phase checking round-to-round consistency.  The main formal design point is that protocol
assembly, completeness, and soundness accounting are expressed through the generic
oracle-reduction composition infrastructure from
Chapter~\ref{chap:oracle_reductions}.

\subsection{Protocol specification}\label{sec:fri_spec}

\begin{definition}[FRI input relation]\label{def:fri_input_relation}
\lean{Fri.Spec.inputRelation}
\uses{def:distance_from_code,def:reed_solomon_code}
    The input relation asserts that the initial oracle statement is $\distance$-close to the
    Reed--Solomon code determined by the initial evaluation domain and degree bound.
\end{definition}

\begin{definition}[FRI output relation]\label{def:fri_output_relation}
\lean{Fri.Spec.outputRelation}
\uses{def:fri_input_relation}
    The output relation is the query-round relation on the final folded statement and witness.
    It is the endpoint of both the folding phase and the full FRI reduction.
\end{definition}

\begin{definition}[Folding-phase protocol]\label{def:fri_folding_protocol}
\lean{Fri.Spec.pSpecFold,Fri.Spec.reductionFold}
\uses{def:fri_input_relation,def:fri_output_relation}
    The folding phase sequentially composes the non-final folding reductions and appends the
    final-fold reduction.  In Lean this is the oracle reduction
    \texttt{Fri.Spec.reductionFold}.
\end{definition}

\begin{definition}[Full FRI reduction]\label{def:fri_reduction}
\lean{Fri.Spec.reduction}
\uses{def:fri_folding_protocol}
    The full FRI reduction appends the query-round oracle reduction to the folding phase.  This
    definitional append is exposed by \texttt{Fri.Spec.reduction\_verifier\_eq\_append} on the
    soundness side.
\end{definition}

\subsection{Perfect completeness}\label{sec:fri_completeness}

Completeness is layered.  Per-round honest-protocol obligations were originally named as
bricks; as of 2026-06-11 every fold-side brick is a proved theorem, and the query-round brick
--- which was \emph{false as stated} (the query verifier checks round consistency for every
round, while the chained relation constrained only the last fold) --- has been repaired and
proved against the full-chain consistency invariant.

\begin{theorem}[Fold-side completeness, discharged]\label{thm:fri_fold_completeness}
\lean{Fri.Spec.Completeness.foldRoundPerfectCompletenessResidual_holds,Fri.Spec.Completeness.finalFoldRoundPerfectCompletenessResidual_holds,Fri.Spec.Completeness.foldPhasePerfectCompletenessResidual_holds,Fri.Spec.Completeness.reductionPerfectCompletenessResidual_holds}
\uses{def:fri_reduction}
    The non-final fold rounds, the final fold round, the assembled folding phase, and the
    reduction-level composition are each perfectly complete; all four former residuals carry
    axiom-clean \texttt{\_holds} theorems, assembled through the generic append and
    sequential-composition keystones.
\end{theorem}

\begin{definition}[Query-round chain invariant]\label{def:fri_query_chain_invariant}
\lean{Fri.Spec.Completeness.queryCheckerAccepts,Fri.Spec.Completeness.queryRoundChainRelation}
\uses{def:fri_output_relation}
    The full-chain consistency invariant, stated in the checker's own currency: the compiled
    query verifier never fails on the statement, at any transcript.  Conjoining it onto the
    base relation makes a relation/check mismatch impossible by design.
\end{definition}

\begin{theorem}[Query-round completeness, repaired]\label{thm:fri_query_completeness}
\lean{Fri.Spec.Completeness.queryRound_perfectCompleteness_repaired}
\uses{def:fri_query_chain_invariant}
    The query-round oracle reduction is perfectly complete from the chain relation to the base
    relation.  The original residual statement
    (\texttt{queryRoundPerfectCompletenessResidual}) is retained, documented as false as
    stated, with a pointer to this repair.
\end{theorem}

\begin{definition}[Chain-delivery residual]\label{def:fri_chain_delivery_residual}
\lean{Fri.Spec.Completeness.queryRoundChainDeliveryResidual}
\uses{thm:fri_fold_completeness,def:fri_query_chain_invariant}
    The single remaining completeness obligation: the folding phase delivers the chain
    invariant (honest folding produces oracles the query checker accepts).  Unlike the
    original query residual this is true by design for honest runs; the proof route is the
    proved fold-phase completeness strengthened on the honest-run support.
\end{definition}

\subsection{Soundness accounting}\label{sec:fri_soundness}

The FRI soundness file now contains both the error-budget definitions and the compositional
theorems tying those budgets to the verifier, conditional on the expected per-round and
per-phase soundness obligations.

\begin{definition}[Per-round and total error]\label{def:fri_error}
\lean{Fri.Spec.roundError,Fri.Spec.queryRoundError,Fri.Spec.queryError,Fri.Spec.totalError}
\uses{def:proximity_gap}
    $\mathsf{roundError}_i$ is the BCIKS20-style proximity-gap error for folding round $i$.
    $\mathsf{queryRoundError}_i$ is the query-consistency contribution for one query check,
    $\mathsf{queryError}$ sums those contributions over all query checks, and
    $\mathsf{totalError}$ adds the fold-phase and query-phase budgets.
\end{definition}

\begin{lemma}[Budget projections]\label{lem:fri_error_projections}
\lean{Fri.Spec.roundError_sum_le_totalError,Fri.Spec.roundError_le_totalError,Fri.Spec.queryError_le_totalError}
\uses{def:fri_error}
    The summed fold-round budget, each individual fold-round budget, and the whole query budget
    are each bounded by $\mathsf{totalError}$.
\end{lemma}

\begin{theorem}[Composed FRI soundness]\label{thm:fri_soundness_total_error}
\lean{Fri.Spec.reduction_soundness_of_phases,Fri.Spec.reduction_soundness_totalError,Fri.Spec.reduction_soundness_totalError_of_rounds}
\uses{def:fri_reduction,def:fri_error}
    If the folding phase is sound with the sum of the per-round errors and the query phase is
    sound with $\mathsf{queryError}$, then the appended FRI verifier is sound with error
    $\mathsf{totalError}$.  The round-level theorem further assembles this from individual
    fold-round soundness facts, final-fold soundness, query-round soundness, and the named
    sequential-composition and append residuals.
\end{theorem}

\begin{remark}[Status of the FRI soundness theorem]\label{rem:fri_soundness_status}
    The theorem above is a real composition theorem about the FRI verifier and the
    $\mathsf{totalError}$ budget.  It is not an unconditional low-degree-testing theorem:
    callers still supply the per-round folding soundness, query-round soundness, and generic
    composition residuals.
\end{remark}

\subsection{Batched FRI}\label{sec:batched_fri}

Batched FRI reduces a batch of polynomial oracles to the base FRI protocol through a batching
round and a virtual oracle lens.  Its security development is where ArkLib connects the FRI
query phase to the correlated-agreement and proximity-gap statements from
Section~\ref{sec:proximity_gap_defs}.

\begin{definition}[Batched FRI reduction]\label{def:batched_fri_reduction}
\lean{BatchedFri.Spec.BatchingRound.batchOracleReduction,BatchedFri.Spec.batchedFRIOracleLens,BatchedFri.Spec.liftedFRI,BatchedFri.Spec.batchedFRIreduction}
\uses{def:fri_reduction,def:reed_solomon_code}
    The single batched round forms a random linear combination of the batch and feeds it to a
    lifted base FRI reduction.  The oracle lens routes the lifted FRI oracle queries back to the
    batch oracles.
\end{definition}

\begin{definition}[Batched FRI structural lens package]\label{def:batched_fri_lens_package}
\lean{Fri.batchedFRIOracleLensReduction,Fri.batchedFRIOracleLensReduction_holds}
\uses{def:batched_fri_reduction}
    The structural part of the batching-to-base-FRI reduction is proved definitionally: the
    constructed lens and lifted reduction have the advertised shape.
\end{definition}

\begin{lemma}[Batched FRI query-round probability and density]\label{lem:batched_fri_query_round_probability}
\lean{Fri.queryRoundAcceptanceBound,Fri.queryRoundAcceptanceBound_holds,Fri.queryRoundAcceptanceBound_of_density,Fri.queryRoundDensityBound,Fri.queryRoundDensityBound_holds,Fri.QueryRound.probabilityAcceptanceBound,Fri.QueryRound.probabilityAcceptanceBound_holds,Fri.QueryRound.probabilityDetectionBound,Fri.QueryRound.probabilityDetectionBound_holds,Fri.Spec.QueryRound.getChallenge_probOutput_uniform_eq_Pr}
\uses{def:batched_fri_lens_package}
    The local query-round analysis is no longer a placeholder.  ArkLib proves both a normalized
    density route, bounding the probability that all repeated queries land in a good set by
    $(1-\delta)^t$, and a probability-space route for the verifier's independent-uniform query
    tuple.  The latter includes the bridge from the concrete FRI query challenge oracle to the
    product PMF used in the standalone probability theorem.
\end{lemma}

\begin{definition}[Batched FRI Claim 8.2 frontier]\label{def:batched_fri_claim82_frontier}
\lean{Fri.fri_query_soundness,Fri.FriQuerySoundnessParts,Fri.fri_query_soundness_of_parts,Fri.fri_query_soundness_of_jointProximity}
\uses{lem:batched_fri_query_round_probability,def:batched_fri_lens_package,lem:joint_proximity_bridges}
    Claim 8.2 of \cite{BCIKS20} is represented as a faithful residual frontier rather than as a
    vacuous theorem.  The split fields isolate the proved query-round bound, the proved
    batching/oracle-lens reduction, and the coding-theoretic bridge from correlated agreement to
    joint agreement.
\end{definition}

\begin{theorem}[Batched FRI Claim 8.2 adapters]\label{thm:batched_fri_query_soundness}
\lean{Fri.fri_query_soundness_of_queryRoundDensityBoundAndBatchedFRIOracleLens,Fri.fri_query_soundness_of_queryRoundProbabilityBoundAndBatchedFRIOracleLens,Fri.fri_query_soundness_of_queryRoundDensityBoundAndBatchedFRIOracleLensAndRSAffineLine,Fri.fri_query_soundness_of_queryRoundProbabilityBoundAndBatchedFRIOracleLensAndRSAffineLine,Fri.fri_query_soundness_of_queryRoundDensityBoundAndBatchedFRIOracleLensAndRSCurve,Fri.fri_query_soundness_of_queryRoundProbabilityBoundAndBatchedFRIOracleLensAndRSCurve}
\uses{def:batched_fri_claim82_frontier,thm:bciks_rs_gap}
    The density and probability adapters reassemble Claim 8.2 once the remaining
    correlated-agreement bridge is supplied.  Specialized affine-line and polynomial-curve front
    doors package the BCIKS20 Reed--Solomon inputs, including the strict coefficient-polynomial,
    boundary, radius, and probability-trigger obligations.
\end{theorem}

\begin{definition}[Batched FRI Claim 8.3 frontier]\label{def:batched_fri_claim83_frontier}
\lean{Fri.fri_soundness,Fri.FriSoundnessParts,Fri.fri_soundness_of_parts,Fri.friSoundnessQueryLift,Fri.friSoundnessSequentialComposition,Fri.friSoundnessTotalErrorAccounting,Fri.friBatchPhaseErrorBound,Fri.friTailPhaseErrorBound}
\uses{thm:batched_fri_query_soundness,def:fri_reduction}
    Claim 8.3 is the end-to-end Batched-FRI soundness residual.  ArkLib splits it into a
    query-soundness lift from Claim 8.2, the sequential composition of the batching round with the
    lifted FRI tail, and total-error accounting for the
    $\epsilon_C + \alpha^\ell$ budget.
\end{definition}

\begin{theorem}[Batched FRI Claim 8.3 assembly]\label{thm:batched_fri_claim83_assembly}
\lean{Fri.batchedFRIreduction_verifier_eq_append,Fri.batchedFRISequentialCompositionSoundness_of_append,Fri.friSoundnessSequentialComposition_of_append,Fri.friSoundnessTotalErrorAccounting_of_phase_bounds,Fri.friSoundnessTotalErrorAccounting_of_named_phase_bounds,Fri.fri_soundness_of_queryRoundDensityBoundAndBatchedFRIOracleLensAndSequentialCompositionAndPhaseErrorBounds}
\uses{def:batched_fri_claim83_frontier}
    The current Claim 8.3 assembly proves the structural append seam, exposes the generic
    sequential-composition theorem for the concrete Batched-FRI reduction, and reduces the
    total-error field to named batching-phase and FRI-tail phase bounds.  The remaining obligations
    are explicit inputs, not hidden assumptions.
\end{theorem}

\begin{theorem}[Small-field batched FRI query soundness]\label{thm:batched_fri_small_field}
\lean{Fri.fri_query_soundness_of_card_le_affineLine,Fri.fri_query_soundness_of_card_le_curve,Fri.fri_soundness_of_card_le_curve}
\uses{thm:batched_fri_query_soundness,thm:bciks_rs_gap}
    In small-field or vacuous regimes, ArkLib discharges the BCIKS20 residual hypotheses using
    the in-tree unconditional correlated-agreement instances, yielding concrete batched FRI
    query-soundness consequences.
\end{theorem}

\begin{remark}[Status of Batched FRI]\label{rem:batched_fri_status}
    The blueprint separates proved protocol plumbing from the still-hard coding-theory fronts.
    Query counting/probability, the structural oracle lens, and several specialized
    Reed--Solomon adapters are proved.  General correlated agreement at the target radius,
    virtual-oracle preservation for every tail context, and the per-phase error bounds needed for
    a fully unconditional Claim 8.3 remain named obligations.
\end{remark}
